### 2024-10-28 manuscript.tex
Updates from Overleaf

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[manuscript,screen,review,acmsmall]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[CSCW '25]{Make sure to enter the correct
  conference title from your rights confirmation emai}{Oct 18--22
  2025}{Bergen, Norway}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Interlocks Mitigate Information Cascades in Large-Scale Social Learning}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Edward L. Platt}
%\authornote{Both authors contributed equally to this research.}
\email{ed@elplatt.com}
\orcid{0000-0003-2148-3841}
\affiliation{%
  \institution{Cinder Oak}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}

\author{Herminio Bodon}
\orcid{0009-0004-3270-3621}
\email{HerminioBodon2020@u.northwestern.edu}
\affiliation{%
  \institution{Northwestern University}
  \city{Evanston}
  \state{Illinois}
  \country{USA}
}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}

\author{Daniel M. Romero}
\orcid{0000-0002-8351-3463}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}}
\email{drom@umich.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Platt, Bodon, \& Romero}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}

\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002950.10003624.10003633</concept_id>
<concept_desc>Mathematics of computing~Graph theory</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003130.10003131.10003570</concept_id>
<concept_desc>Human-centered computing~Computer supported cooperative work</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010341.10010349.10010355</concept_id>
<concept_desc>Computing methodologies~Agent / discrete models</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Mathematics of computing~Graph theory}
\ccsdesc[500]{Human-centered computing~Computer supported cooperative work}
\ccsdesc[300]{Computing methodologies~Agent / discrete models}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Collaboration, Social Learning, Networks, Interlocks, Information Cascades, Polarization, Deliberation, Consensus}

\received{16 January 2024}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
Large collaborations face a dilemma:
participatory governance is both
highly desirable and notorioulsy difficult
to implement in large groups.
Elinor Ostrom's empircal study of commons governance identified participatory rule-making
as one of the key principles necessary for the success and longevity of resource regimes
\cite{ostrom_collective_2000, ostrom_governing_1990}.
Participatory governance, particularly participatory deliberation,
enables the aggregation and synthesis of diverse knowledge
\cite{
dewey_creative_1940,
anderson_epistemology_2006,
ackerman_deliberation_2002},
benefits from the wisdom of crowds
\cite{
surowiecki_wisdom_2005,
hill_group_1982,
hong_groups_2004,
golub_naive_2010,
galton_vox_1907},
identifies and resolves conflicts
\cite{gentry_consensus_1982},
and incentivizes future cooperation by increasing perceptions of trust and fairnesss
\cite{
ostrom_collective_2000,
bowles_endogenous_1998}.
Unfortunately, as groups grow larger so does the difficulty of participatory governance.
The challenges of participatory governance in large groups include:
increased time and effort needed to reach decisions
\cite{
fishkin_voice_1997,
gentry_consensus_1982,
steiner_group_1972},
the emergence of power inequalities
\cite{
freeman_tyranny_1972,
shaw_laboratories_2014,
kittur_power_2007,
boehm_egalitarian_1993},
and counterproductive social dynamics such as
information cascades \cite{banerjee_simple_1992},
polarization \cite{schkade_what_2007},
social loafing \cite{karau_social_1993},
and the majority illusion \cite{lerman_majority_2015}.
Yet examples of successful large-scale participatory governance do exist, including Wikipedia \cite{giles_internet_2005, keegan_evolution_2017, forte_scaling_2008}, free and open source software projects \cite{benkler_coases_2002}, grassroots protest and crisis-response networks \cite{manilov_occupy_2013, tufekci_twitter_2017,
gonzalez-bailon_networked_2016,
landau_place-based_2017,
brugh_combining_2019}
and self-managed organizations
\cite{laloux_reinventing_2014,monsen_buurtzorg_2013,robertson_holacracy_2015}.
These examples often rely on the unprecedented fast, inexpensive, many-to-many communication afforded by the internet.
But not all intenet-enabled participatory collaborations succeed \cite{tufekci_twitter_2017, geiger_does_2009}. Why do some succeed while others fail?

In this paper, we begin with a review of the literature on large-scale participatory decision-making and related topics.
We focus in particular on empirical studies of successful large-scale participatory projects, looking for commonalities.
We then focus on one such commonality: social network structure,
specifically interlock networks.
We also narrow our focus to social learning:
the process by which individuals revise their beliefs and opinions
based on obersvations and interactions with others---a
core part of participatory decision-making.
Finally, we develop an agent-based model
to test the effects of interlock networks on social learning
in large groups.
Our findings suggest that interlock network structure may contribute to the success of large-scale participatory projects by providing protection from information cascades and polarization.

\subsection{Empirically-Motivated Simulation}

If large-scale participatory decision-making is difficult,
then studying it is doubly so.
The large number of individuals involved makes lab experiments prohibitively expensive.
Randomized, controlled field experiments would require dividing
a community into control and treatment arms
and ensuring that the members of each
arm do not interact,
which is both impractical and potentially
against the best interests of the group members,
especially if they are making consequential decisions.
For the present study, we use computer simulation models to help bridge the gap
between empirical observation and future experimentation.

Readers may understandably question how well a computer simulation can provide insight into the rich complexities of human behavior.
As with all models, computer and otherwise,
the usefulness of our methodology depends on the soundness of our simplifying assumptions and the applicability of our outcome measures.
So we base both on a thorough review of the observational literature.
If our model finds that the conditions present in real-world participatory groups lead to the outcomes observed in real-world participatory groups,
we can provide insight on how and why those conditions might lead to those outcomes.

Specifically, we use Agent-Based Modeling (ABM),
well-established type of simulation model often used in complex systems,
economics, and biology
\cite{hong_groups_2004,
lazer_network_2007,
golub_naive_2010,
zollman_social_2012,
grim_scientific_2013,
barkoczi_social_2016,
gomez_clustering_2019}.
Simply put, we simulate a large number of agents, each modeling a single group participant.
We allow those agents to interact based on well-defined behavioral strategies,
and observe the progression of their states over time.
We describe our model in greater detail in
Sections \ref{subsec:abm}---\ref{subsec:nk} and \ref{sec:methods}.
Our empirically-motivated simulation approach allows us to isolate individual factors and perform controlled experiments with a precision and scale not feasible in a real-world setting, while remaining grounded in real-world observations.
Our findings can establish the plausibility of empirically-motivated hypotheses,
provide insight into the underlying mechanisms,
and provide guidance and motivation for future real-world empirical studies.

\subsection{Collaboration and Network Structure}
As we will elaborate below,
we find that real-world examples of successful large-scale participatory collaborations
often share a common trait:
interlocking social network structure.
Here, we use ``network'' in the sense of a set of entities with dyadic relationships between them \cite{newman_networks_2018}, e.g.,
individuals linked to other individuals through direct communication.
Similarly, we use ``structure'' to refer to the configuration of individual dyadic relationships and connnections rather than in the ``social system'' sense, e.g., the difference in communication links between a roundtable discussion and a lecture.
An {\em interlocking network} or {\em interlock}
is a network in which groups of entities are connected (interlocked)
through shared individual constituents \cite{taylor_world_2015}.
As a much-studied example,
an interlocking directorate is a network of corporations
in which two corporations (the groups) are considered connected if any individual director (the constituent) belongs to both of their boards 
\cite{mizruchi_what_1996}.
Or as a more modern example,
collaborative interactions on Wikipedia take place on individual articles, their per-article talk pages, and on topical WikiProject pages \cite{platt_network_2018}.
In this example, articles/talk pages/projects are the groups, and individual editors are the constituents.

The relationship between network structure and large-scale participatory collaboration may not be immediately obvious.
In large collaborations, while it may be the case than anyone {\em can} talk to anyone else, it is not the case that everyone {\em does} talk to everyone else.
The number of potential interactions is much greater than the capacity of the collaborators.
This fact implies a network structure: who interacts with whom.
While hierarchical ``tree'' newtorks are a practical way to interconnect a large group of people,
as in a military chain of command or corporate org chart,
they privilege some group members above others.
Participatory collaboration requires different structures.
In the context of self-organized collaboration, most network-based analyses have focused on 
small-world networks \cite{watts_collective_1998}
and preferential attachment networks \cite{barabasi_emergence_1999},
rather than interlocks,
with the notable exception of Salehi \& Bernstein's work on network rotation \cite{salehi_hive:_2018}.

There is considerable evidence that group composition
\cite{hill_group_1982,
nishi_inequality_2015,
robert_jr_differences_2018,
sydow_diversity_2017,
lerner_diverse_2018,
arazy_information_2011}
and social network structure
\cite{
kearns_experiments_2012,
barkoczi_social_2016,
salehi_hive:_2018,
mason_propagation_2008,
mason_collaborative_2012}
can influence the success of a collaboration.
Interlocking structure has implications for both of these factors.
By dividing an entire collaborative project into smaller groups,
interlocks influence group size and group composition,
potentially sidestepping some of the pitfalls of large group collaboration.
Similarly, by creating interconnections between groups through mutual membership,
interlocks influence the overall strucutral properties of the network.
In turn, those structural properties influence the speed and flow of information diffusion through the entire population of collaborators.

\subsection{Social Learning}
To begin to disentangle the multiple ways interlocking structure might influence the large participatory collaborations,
we focus on a subset of possible factors.
Specifically, we model collaborations as {\em social learning} processes \cite{barkoczi_social_2016, mason_collaborative_2012, hong_interpreted_2009, degroot_reaching_1974},
in which homogeneous agents exchange information and choose preferred solutions to a problem based on a combination of individual assessments of solution quality and information received from peers.
In this model, agents can apply different
{\em social learning strategies} to integrate socially acquired information with their own individual assesments.
Strategies can be classified as {\em conformist} when social influence is prioritized over individual assesments and {\em individualist} in the opposite case.
While this model leaves out many real-world collaboarative considerations,
ranging from the deep
(e.g.,
argumentation \cite{fishkin_voice_1997, mansbridge_minimalist_2015},
identity \cite{bail_breaking_2022,schkade_what_2007},
implicit bias \cite{amodio_stereotyping_2006},
interpersonal conflict\cite{whiting_did_2019})
to the mundane (e.g.,
mood, 
conversational tone),
solution-quality-based considerations are common across collaborations,
and are considered by many scholars to be a normative or even necessary
part of participatory decision-making
\cite{mansbridge_minimalist_2015}.
We thus propose this work as an starting point for understanding the seemingly-important role of interlocks in large-scale participatory collaboration, leaving additional considerations as crucial future work.

\subsection{Research Questions}
Our primary focus is whether interlocking network structure improves the quality of social-learning-based collaboration:
\begin{quote}
{\bfseries RQ 1.} Does interlocking structure influence solution quality in large-scale social learning?
\end{quote}

As higher quality collaborative output has sometimes been observed to require longer spans of time
\cite{
gentry_consensus_1982,
platt_network_2018},
we also ask whether interlocking structure
has any such influence:

\begin{quote}
{\bfseries RQ 2.} Does interlocking structure influence speed of convergence in large-scale social learning?
\end{quote}

Social learning strategies have been shown to modulate the influence of network structure on collaborations
\cite{barkoczi_social_2016, zollman_social_2012},
so we also ask whether the same is true for the effects of interlocking networks:

\begin{quote}
{\bfseries RQ 3.} Are the effects of interlocking structure on social learning modulated by social learning strategy?
\end{quote}

Finally, as polarization is one of the mechanisms known to hinder quality in social learning and collaboration
\cite{schkade_what_2007, zollman_social_2012},
we ask whether interlocking network strucutre has an effect on polarization:

\begin{quote}
{\bfseries RQ 4.} Does interlocking structure influence polarization in large-scale social-learning?
\end{quote}

\subsection{Contributions}
In this paper, we present the following contributions:
\begin{itemize}
\item{We show that social-learning on interlocks outperforms small-world and preferential attachment networks for some social learning strategies, and performs comparably for others (RQ 1).}
\item{Specifically, we show that in groups using a conformist social learning strategy, interlocks outperform other networks (RQ 3).
We attribute this improvement to a protective effect against information cascades that lead to the adoption of popular but sub-optimal solutions.}
\item{We also show that interlocks do not result in notably slower convergence, even in cases where they improve solution quality (RQ 2).}
\item{Finally, we show that interlocks reduce polarization when agents are using a conformist strategy (RQ 4).
Again, we attribute this result to a protective effect against information cascades.}
\end{itemize}
Taken together, our findings show that interlocks provide a protective effect against information cascades in large-scale social learning, which improves the performance of conformist strategies while resulting in similar performance for individualist strategies.
This protective effect may contribute to the relative success observed in large participatory collaborations with interlocking network structure.

\section{Background}

Examples of large participatory collaborations have become increasingly common in recent years,
and some have achieved surprising levels of success.
In particular, many of these recent successes rely on the fast, inexpensive, many-to-many communication afforded by the internet,
which has enabled an environment Jochai Benkler calls the ``networked information economy''
\cite{benkler_wealth_2006}.
Wikipedia, the popular online encyclopedia has been produced by millions of editors (47.6 million for just the English language edition at time of writing \cite{noauthor_list_2024, noauthor_wikipediawikipedians_2024})
using consensus-based, participatory decision-making \cite{keegan_evolution_2017, forte_scaling_2008}.
Despite common wisdom regarding the impossibility of large-scale participatory governance,
Wikipedia has been found comparable in quality to traditionally-edited encyclopedias \cite{giles_internet_2005},
leading to the popular Wikipedian in-joke that
``The problem with Wikipedia, is that it only works in practice. In theory, it's a total disaster'' \cite{elsharbaty_editing_2016}.
Similarly, Free and Open-Source Software (FOSS),
which is also highly participatory,
has achieved a level of success that rivals traditionally-produced software
\cite{benkler_wealth_2006, raymond_cathedral_1999}.
As one example, the FOSS operating system Linux is the most popular OS for internet servers and the core of the Android OS for mobile devices.
In a business context, examples of participatory worker self-management can be found in both corporate and cooperative contexts \cite{laloux_reinventing_2014, lee_self-managing_2017}.
Examples include Zappos.com's evolving use of the Holacracy system
\cite{robertson_holacracy_2015, groth_zappos_2020}
and the Buurtzorg model of homecare
\cite{monsen_buurtzorg_2013, gray_home_2015}.
Grassroots social movements and crisis-response groups have also made use of participatory governance.
The Occupy movment embraced participatory governance \cite{detar_intertwinkles:_2013, gonzalez-bailon_networked_2016}
which resurfaced in the 60,000 self-organized volunteers of the Occupy Sandy crisis-response efforts \cite{landau_place-based_2017, brugh_combining_2019}.
Similarly, participatory practices were a key part of the 2013 Gezi Park protests
\cite{tufekci_twitter_2017}.

\subsection{Interlocks}
When reviewing these examples, a remarkably common pattern stands out: interlocking network strucutre.
In these successful examples of large-scale participatory projects,
communication is frequently organized into small participatory groups of peers, based on interest, region, function, or simply happenstance.
These groups are then interconnected through individuals who participate in multiple groups.
We reproduce excerpts from several examples here to demonstrate the ubiquity of this pattern.

Occupy Wall Street and Occupy Sandy \cite{manilov_occupy_2013}:
\begin{quote}
One of the greatest strengths of the Occupy movement has been its willingness to decentralize decision making into networks built on a foundation of small groups that move easily between online and real-world activities.
\end{quote}

The 2013 Gezi Park protests in Turkey \cite{tufekci_twitter_2017}:
\begin{quote}
The next day, the protesters at the park initiated a ``small-forum'' process,
inviting people to break into small groups to discuss the government’s proposal. 
\end{quote}

Blogs on the early web \cite{benkler_wealth_2006}:
\begin{quote}
Filtering, accreditation, synthesis, and salience are created through a system
of peer review by information affinity groups, topical or interest based.
These groups filter the observations and opinions of an enormous range of people,
and transmit those that pass local peer review to broader groups and ultimately to the polity more broadly...
\end{quote}

US anti-war protests \cite{detar_intertwinkles:_2013}:
\begin{quote}
The core of these protests were
organized by clusters of small-scale affinity groups. Each affinity
group, consisting of 10 to 20 people, decided what sort of role they
were comfortable fulfilling, and then coordinated with other affinity
groups to participate in larger actions.
\end{quote}

Collaborative design \cite{salehi_hive:_2018}:
\begin{quote}
We argue that large
collectives cannot cross-pollinate ideas simply by placing all ideas into the commons for others to
see — people in the collective must instead interact with each other closely to benefit from ideas and
perspectives that each bring to the table.
We define an approach, network rotation, that supports
this goal by gradually weaving together team membership from across the collective to increase
the efficiency of the underlying collaboration network while maintaining high tie strength within
teams.
\end{quote}

The above examples all share the same social network structure of small groups connected through mutual membership: {\em interlocks}.
Interlocks have been well-studied in several other contexts.
Perhaps the most well-known is the study of intelocking directorates,
the practice of one individual sitting on the board of directors for
multiple corporations,
allowing an implicit flow of information and influence
\cite{sapinski_interlocking_2018}.
Interlocks arise in other fields as well,
including revolutionary cells
\cite{fishcer_paul_1994, healey_using_2013},
the UK suffragette movement \cite{crossley_covert_2012},
and the global network of cities \cite{taylor_world_2015}.

\subsection{Participatory Governance and Social Learning}
This paper touches on several related topics, namely: problem-solving, decision-making, deliberation, and social learning.
These concepts are very closely related, but not equivalent,
potentially leading to confusion.
Matters are made worse by differing (and conflicting) definitions in the literature.
Here, we attempt to untangle this conceptual knot.

The motivation and ultimate focus of our work is {\em participatory governance}
in the sense Elinor Ostrom's third design principle for long-suriving resource regimes: that ``most of the individuals affected by a resource
regime can participate in making and modifying their rules.''
\cite{ostrom_collective_2000}
In this sense, governance is largely about {\em decision-making} regarding
rules and their enforcement.
But we are careful to note that rules and decisions can implicit
{\em rules-in-use} in addition to explicit {\em rules-in-form}
\cite{hess_framework_2005, keegan_evolution_2017}.

\subsubsection{Deliberation}
{\em Deliberation} is an exemplar of participatory governance.
But prior to the internet and other forms of digital communication,
the logistics of large-scale governance generally excluded more participatory
forms of decision-making \cite{ackerman_deliberation_2002}.
More tractable alternatives such as voting and representative
assemblies are less participatory.
There is no consensus on the exact definition of deliberation.
We will follow the ``minimalist'' definition proposed by Dryzek
\cite{dryzek_deliberative_2002} and refined by Mansbridge \cite{mansbridge_minimalist_2015}:
\begin{quote}
Deliberation is mutual communication that involves weighing and reflecting on preferences, values and interests regarding matters of common concern.
\end{quote}
Note that this definition does not actually require the making of a decision, but by explicitly mentioning ``preferences," does imply the existence of a decision to be considered.
And again, while ``deliberation'' may conjure images of formal institutions,
note that the above definition includes both formal and informal communication,
as long as it is mutual and non-coercive.

\subsubsection{Social Learning\label{subsec:social}}
In {\em social learning}, individuals revise their beliefs by observing and interacting with others.
Social learning allows an individual to augment their limited information and perspective by observing others who may have different information
and perspectives \cite{hong_interpreted_2009, golub_naive_2010}.
When this observation takes the form of mutual communication,
social learning fits nicely into the above definition of deliberation.
In fact, many models of social learning implicitly assume mutual communication
\cite{golub_naive_2010, lazer_network_2007, mason_collaborative_2012, barkoczi_social_2016, gomez_clustering_2019}.
Deliberation is a thus a type of social learning,
and the results from the above models should be expected to apply.

Why do people engage in social learning?
Usually to solve problems.
Specifically, to solve problems too complex or resource-intensive to
solve individually.
In other words, social learning is also a form of collective
problem-solving and collective intelligence.
Participatory governance is such a problem.

In the case of participatory governance,
there are two related, but subtly different problems being solved.
The first problem is for participants try to accurately determine which outcome would maximize social utility.
For example, a group deciding what type of pizza to order might gather information on other participants' most and leaast favorite pizza toppings.
The second problem is for participants to
advocate for their desired outcome,
which might be based on their personal preference
or on the expected social utility,
e.g., arguing for their favorite pizza topping
or for a pizza topping they believe will satisfy the most people.
The former problem is a matter of information and mathematics,
while the latter is a matter of taste and rhetoric.
Matters of taste are beyond our methods, so our models focus on the former problem of gathering information and maximizing utility.
So we model a group's ability to identify outcomes that maximize
common good.
This is a necessary, but not sufficient condition for the group to actually
adopt such an outcome.

\subsubsection{Putting It All Together}
Summarizing the above, we use an agent-based model to simulate a group's
ability to collectively problem-solve
---
through the use of mutually-communicative,
deliberation-like social learning
---
the problem of fidning a governance outcome that maximizes the social utility of the group.
To the extent that a group prioritizes such outcomes,
our models should provide meaningful insight.

\subsection{When Social Learning Goes Wrong}
While social learning can allow members of a group to combine their knowledge and resources,
it can also introduce new problems.
When information is repeated without sufficient independent checking,
a false belief starting with one person can be repeated and amplified until it is accepted by a large number of people:
an {\em information cascade} \cite{banerjee_simple_1992}.
Similarly, in the {\em majority illusion},
an individual may incorrectly infer that the majority of a group holds a particular belief if they are only exposed to others who do hold that belief \cite{lerman_majority_2015}.
Finally,
rather than reaching a single consensus,
social learning can cause {\polarization},
in which a group fragments into two or more factions which hold similar beliefs within each group,
but conflicting beliefs between groups
\cite{schkade_what_2007}.
In groups that successfully implement large-scale participatory decision-making,
these pitfalls must be avoided.
One of this paper's main goals is to identify potential structural or behavioral mechanisms that might allow a group to do so.

\subsection{Agent-Based Modeling\label{subsec:abm}}
Our Agent-Based Model (ABM) closely follows previous ABM work on social learning
\cite{lazer_network_2007, barkoczi_social_2016, gomez_clustering_2019}.
The ABM consists of a large collection of {\em agents}:
simulated entities, each with their own private information and the ability to update that information based on interactions with other agents.
In our case, the agents are intended to represent
individual participants in a participatory decision-making
process.
The exact behavior of the agents is determined by three factors:
\begin{description}
\item[Objective Function]
At the beginning of each ABM run, an objective function $Q(s)$ is defined.
The goal of the agents is to find an $s$ that maximizes
$Q(s)$.
\item[Social Network]
The social network provides links between some pairs of agents and not others.
These links allow neighboring agents to interact with each other and share information.
\item[Social Learning Strategy]
The social learning strategy is a simple set of rules the agents use to interact with their neighbors and update their belief about the best value of $s$.
\end{description}

At the beginning of a simulation run, each agent is initialized with a random $s$.
The run proceeds in discrete time steps.
At each time steps, each agent interacts with each of its neighbors according to its strategy and
updates its belief.
After a fixed number of time steps, the run ends.
Ideally, the number of time steps is large enough that
the agents' beliefs have reached a stable state.
Different agents may still have different beliefs,
but at this point the distributions of beliefs and the potential outcomes of votes can be analyzed.
This process is repeated for many runs,
each with a new set of initial conditions and objective $Q(s)$.
By performing many runs, statistical differences between different networks and/or strategies can be identified.

\subsection{NK Model\label{subsec:nk}}
For our agent-based model, we need an objective function $Q(s)$ for the agents to maximize.
We follow previous literature on social learning
\cite{lazer_network_2007,barkoczi_social_2016,gomez_clustering_2019}
by using the NK Model \cite{kauffman_towards_1987,weinberger_local_1991} to create our objective function.
The NK Model has a number of properties that make it a good candidate for modeling social learning.
It was originally developed to model the natural selection of genetic mutations \cite{kauffman_towards_1987},
but we describe it below in terms of a more relevant example.

Let's consider a referendum on $N$ different yes/no proposals.
These proposals might, for example, be of the following sort:
\begin{itemize}
    \item{Should a new housing development be approved at 123 Main Street?}
    \item{Should electric scooters be permitted on sidewalks?}
    \item{Should a traffic lane be added to Anytown Highway?}
\end{itemize}

Individuals may feel strongly about some of the proposals and not at all about others.
Rather than considering individual voters, let's consider voting blocs:
groups of people who are all concerned about the same issues.
Specifically, let's assume that there are $N$ voting blocs, each mostly concerned with one of the $N$ proposals.
However, the result of implemenmting any single proposal might depend on which of the other propsals are implemented.
One bloc for example, let's call them the {\em Bloc for Beds, Bikes, and Busses}, might support approving new a housing development,
but only if additional public transportation and bike lanes are created.
So we also assume that each bloc is concerned with $K$ additional proposals, for a total of $K+1$.

After the vote is held and the results have been implemented, each bloc $n$ will have some approval rating $q_n(s)$,
where $s$ is an $N$-bit binary number with each bit representing the outcome of one yes/no proposal.
Each bloc's approval rating will depend on the outcomes of the $K+1$ relevant proposals.
Note that we don't need to assume that the voters themselves know or can predict ahead of time what their approval rating for each possible outcome would be.
By averaging the approval ratings of each bloc for a given outcome $s$, we get a number representing the social utility, fitness, or {\em quality} of that solution:
\begin{eqnarray}
    Q(s) &=& \frac{1}{N}\sum_{n=1}^N q_n(s).
\end{eqnarray}
Any such function $Q(s)$ is an instance of the NK Model.

In this paper, as is typical with the NK Model, we determine the $q_n$ randomly.
First by chosing the relevant $K$ bits of the solution at random, and then by assigning a uniform random value in [0,1] $q_n$ for each possible input.
Random $q_n$ are suitable when we care more about how agents incorporate information about preferences than about why those preferences exist in the first place.

\subsubsection{The NK Model in Participatory Decision-Making}
The above description uses the language of proposals, voting blocs, and approval ratings. This language should not necessarily be taken literally.
We have used it to motivate the NK Model as a suitable model for complex policy decisions.
However, the NK model is just as applicable to a situation where, for example, the "voting blocs" are farm crops, the "policies" are farming practices, and the "fitness" is the crop yield; or to genes, mutations, and reproduction rates, as in the original conception of the NK Model \cite{kauffman_towards_1987}.

We draw attention to the above distinction because of a subtle, but very important point.
As discussed in Section \ref{subsec:social} there are two related but separate approaches to participatory governance.
One, to advocate for a preferred outcome based on preferences;
the other, to determine which outcome would best satisfy everyone's preferences.
The NK Model bridges these two approaches.
It asks: given a random set of subjective preferences,
what is the objective quality of a particular outcome?

\subsubsection{Properties of the NK Model}
The real-world applicability of our findings will depend on how well a real-world system agrees with the assumptions of our model.
We enumerate key assumptions and properties below.
\begin{description}
\item[Discrete solution space]
The NK model solution space consists of discrete states.
This space best represents a set of yes/no questions.
However, the NK model can also be applied to any solution space that can be approximated by a discrete set of solutions, for example, by dividing a continuous range into bins.
\item[Large solution space]
There are $2^N$ possible solutions for any particular instance of the NK model.
The number of solutions grows exponentially with the number of bits in $s$.
The NK model is thus well-suited to representing problems with far more potential solutions than could ever be evaluated.
For example, for N=33, there would be more potential solutions than humans alive on Earth (at time of writing).
\item[Rugged solution space]
Often called a ``rugged landscape,'' the NK model has many local maxima and minima,
often described as ``hills'' and ``valleys.''
A simple hill-climbing search by changing one solution bit at a time will lead to a local maximum,
but there is no way to know if a better solution exists, other than searching for it.
This is a common (if frustrating) feature of real-world problems.
\item[Tunably rugged]
In addition to being rugged, the ruggedness of the NK model can be ``tuned'' by changing the $K$ parameter.
For a value of $K=0$, each solution bit can be optimized independently, and there is a single local maximum.
For the maximum value of $K=N-1$,
the landscape is essentially random.
For intermediate values of $K$,
there are a large number of local maxima, but they are surrounded by ``basins of attraction'' which allow those maxima to be found from nearby solutions using hill-climbing.
These intermediate values are typically the ones used to mimic real-world rugged landscapes.
\item[NP-complete]
For $K \geq 3$, the problem of finding the global optimum of the NK model in general is NP-complete \cite{weinberger_np_1996}.
This means that the NK model is computationally equivalent to the hardest problems solvable by a computer,
and can thus model the complexity of any problem that is capable of being modeled computationally.
\end{description}

\subsection{Social Learning Strategies}

\subsection{Network Types}

TODO

\section{Methods\label{sec:methods}}
Agent-based model

\section{Results}

\section{Discussion}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}

\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\section{Online Resources}

\end{document}